Microlensing as a possible probe of event-horizon structure in quasars

### 2016-09-22 paper1/ms.tex
Corrections
other corrections

\section{introduction}

Active galactic nuclei are thought to be powered by the accretion of
matter from the proximal environment into a supermassive black hole.
The radiation emitted excites the surrounding medium which becomes
detectable as narrow line regions, broad line regions and optical
continuum.  In the direction perpendicular to the accretion disc,
where the medium is more transparent, the outflow tends to be
collimated to jets.  If the orientation is such that the observer's
view of the central region is not blocked by the accretion disc or by
a jet, a quasar is seen \citep[e.g.,][]{1984RvMP...56..255B}.  While
the basic mechanism \citep[originating in the work
  of][]{1964ApJ...140..796S,1964SPhD....9..246Z,1969Natur.223..690L}
is not in doubt, the central engines, near the event horizons of the
black holes, remain to be probed.

@@

In the present paper, we consider a possible indirect method, related
to \cite{1999ApJ...524...49A}, through which gravitational
microlensing could probe the black hole shadow and its proximal quasar
environment.

@@

A simple example of a caustic network is shown in
Figure~\ref{fig:magnification_map}.  There are two general categories
of caustics in gravitational lensing, cusps and fold, and examples of
both kinds can be seen in this figure.  Magnification near a caustic
has universal properties, independent of the system and has been
extensively studied
\citep{1986ApJ...310..568B,1992A&A...260....1S,2002ApJ...574..970G,2002ApJ...580..468G}.
In particular, at distance $p$ from a fold caustic
\begin {equation}
 \mu(p) = \mu_0 + C_0 \frac{1}{\sqrt{p}} \Theta(p).
\end {equation}
Here the magnification of a point source near a caustic is equal to
the sum of the magnification due to other reasons $\mu_0$, assumed to
be locally constant, and a decrease with the square root of the
distance from the fold. The latter term becomes activated only after
the source enters the region interior to the caustic curves when the
values of the step function $\Theta(p)$ become unity. The
proportionality constant $C_0$ depends on the local conditions in the
vicinity of the caustic.

@@

The function $p_s(t)$ can be chosen to be equal to $v_p(t-t_0) + p_{s0}$. Where $p_{s0}$ is the coordinate $p$ of the source at the initial time, and $v_p$ is the component of the velocity
along the $p$ axis. Such a modelling of the motion of the object in the source plane describes a linear motion with constant velocity. Furthermore, we reduce the complexity of the model by 
choosing the function $a(t)$ to be constant in time.

@@

There are four characteristic points visible on the resulting one-dimensional light profile of the crescent shaped source. Two of the points, p1 and p4, mark the outer boundaries of the luminous disc component. The other two, p2 and p3, mark the boundaries of the dark disc component. Since $S_{1D}$ is a projection of $S_{2D}$ on a line perpendicular to the caustic, the following relation holds:
\begin{equation}
    p_2-p_1 = R_p -R_n - a.
\end{equation}
In addition there are two other obvious relations the characteristic points which are independent of the projection:
\begin{equation}
    p_4 -p_1 = 2 R_p,
\end{equation}

@@

\begin{equation}
    R_p = \frac{v_p \left( t_4 -t_1 \right)}{2}, 
\end{equation}

\begin{equation}
        R_n = \frac{v_p \left( t_3 -t_2 \right)}{2}, 
\end{equation}

\begin{equation}
        a = \frac{v_p \left( t_4 +t_1 - t_3 - t_2 \right)}{2}. 
\end{equation}

@@

Figure 4 reveals that the lightcurve of a crescent source has more
visible features than the other two light-curves corresponding to the
disc and gaussian shape. There are three regions where $S_{1D}$ and by
inheritance, the lightcurve has distinguishable behaviour. The first
region that would be recorded on a lightcurve plot represents the
period of time when the bright disk begins to be overlapped by the
caustic and stops when just before the dark disk reaches the
caustic. During this period of time, the flux of light from the source
is increasingly magnified. The second period starts and ends with the
overlapping of the dark disc. As a boundary of the two regions, there
is a distinguishable point where the slope of the magnification is
drastically changed. This apparent discontinuity in the first
derivative of the magnification function is caused by the caustic
amplification of the sudden drop in the $S_{1d}$ function. During the
respective period, the magnification growth slows or even reverses
during the first part of the period and starts to grow faster again as
the dark disk ends its overlap with the caustic. At the point where
the dark disc clears the caustic, the growth of the magnification is
infinite, which appears as a saddle point on the lightcurve. Next, the
final period corresponds to the case when the dark disc has cleared
the caustic and the bright disc continues to overlap with the
caustic. During this period, the lightcurve reaches a peak that for
most of the parameter space is global and for the rest of the
parameter space local.  At the end of the period, the growth of the
magnification is negative infinite. the respective point appears as a
second saddle point on the lightcurve. Past this point the
magnification of any finite source will decay in roughly the same
manner.

@@

\begin{figure}
\centering
    \includegraphics[width = 0.48\textwidth]{figures/M87_shape.eps}
        \includegraphics[width = 0.48\textwidth]{figures/M87_lc.eps}
\caption{\label{fig:M87_plots} The one dimensional profile obtained
  from the numerical integration along the ordonata axis of the source
  image is presented in the upper panel. The corresponding
  lightcurve associated to the 1D profile is presented in the lower
  panel. The points p2 and p3 mark the start and end of the equivalent
  dark disk and are associated with the moments of their overlap with
  the caustic: t2 and t3.}
\end{figure}

>>>>

\section{The General Picture }

Active galactic nuclei (AGN) are thought to be powered by the accretion of
matter from the proximal environment into a supermassive black hole.
The radiation emitted excites the surrounding medium which becomes
detectable as narrow line regions, broad line regions and optical
continuum.  In the direction perpendicular to the accretion disc,
where the medium is more transparent, the outflow tends to be
collimated to jets.  If the orientation is such that the observer's
view of the central region is not blocked by the accretion disc or by
a jet, a quasar is seen \citep[e.g.,][]{1984RvMP...56..255B}.  While
the basic mechanism \citep[originating in the work
  of][]{1964ApJ...140..796S,1964SPhD....9..246Z,1969Natur.223..690L}
is not in doubt, the central engines, near the event horizons of the
black holes, remain to be probed.

@@

In the present paper, we consider a possible indirect method, related
to \cite{1999ApJ...524...49A} and \cite{2015ApJ...814L..26M}, 
through which gravitational
microlensing could probe the black hole shadow and its proximal quasar
environment.

@@

A simple example of a caustic network is shown in
Figure~\ref{fig:magnification_map}.  There are two general categories
of caustics in gravitational lensing, cusps and fold, and examples of
both kinds can be seen in this figure.  Magnification near a caustic
has universal properties, independent of the system and has been
extensively studied
\citep{1986ApJ...310..568B,1992A&A...260....1S,2002ApJ...574..970G,2002ApJ...580..468G}.
In particular, at distance $p$ from a fold caustic
\begin {equation}
 \mu(p) = \mu_0 + C_0 \frac{1}{\sqrt{p}} \Theta(p).
\end {equation}
Here the magnification of a point source near a caustic is equal to
the sum of the magnification due to other reasons $\mu_0$ 
(such as zero point magnification and the effect of other weaker and 
distant lenses), assumed to
be locally constant, and a decrease with the square root of the
distance from the fold. The latter term becomes activated only after
the source enters the region interior to the caustic curves when the
values of the step function $\Theta(p)$ become unity. The
proportionality constant $C_0$ depends on the local conditions in the
vicinity of the caustic.

@@

The function $p_s(t)$ can be chosen to be equal to $v_p(t-t_0) + p_{s0}$. Where $p_{s0}$ is the coordinate $p$ of the source at the initial time, and $v_p$ is the component of the velocity
along the $p$ axis. Such a modelling of the motion of the object in the source plane describes a linear motion with constant velocity. Furthermore, we reduce the complexity of the model by 
choosing the function $a$ to be constant in time.

@@

In figure \ref{fig:char_points} There are four characteristic points visible on the resulting one-dimensional light profile of the crescent shaped source. Two of the points, p1 and p4, mark the outer boundaries of the luminous disc component. The other two, p2 and p3, mark the boundaries of the dark disc component. Since $S_{1D}$ is a projection of $S_{2D}$ on a line perpendicular to the caustic, the following relation holds:
\begin{equation}
    p_2-p_1 = R_p -R_n - a.
\end{equation}
In addition there are two other obvious relations the characteristic points which are independent of the projection:
\begin{equation}
    p_4 -p_1 = 2 R_p,
\end{equation}

@@

\begin{equation}
    R_p = \frac{v_p \times  \left( t_4 -t_1 \right)}{2}, 
\end{equation}

\begin{equation}
        R_n = \frac{v_p \times \left( t_3 -t_2 \right)}{2}, 
\end{equation}

\begin{equation}
        a = \frac{v_p \times \left( t_4 +t_1 - t_3 - t_2 \right)}{2}. 
\end{equation}

@@

Figure \ref{fig:lightcurve_gauss} reveals that the 
lightcurve of a crescent source has more
visible features than the other two light-curves corresponding to the
disc and gaussian shape. There are three regions where $S_{1D}$ and by
inheritance, the lightcurve has distinguishable behaviour. The first
region that would be recorded on a lightcurve plot represents the
period of time when the bright disk begins to be overlapped by the
caustic and stops when just before the dark disk reaches the
caustic. During this period of time, the flux of light from the source
is increasingly magnified. The second period starts and ends with the
overlapping of the dark disc. As a boundary of the two regions, there
is a distinguishable point where the slope of the magnification is
drastically changed. This apparent discontinuity in the first
derivative of the magnification function is caused by the caustic
amplification of the sudden drop in the $S_{1d}$ function. During the
respective period, the magnification growth slows or even reverses
during the first part of the period and starts to grow faster again as
the dark disk ends its overlap with the caustic. At the point where
the dark disc clears the caustic, the growth of the magnification is
infinite, which appears as a saddle point on the lightcurve. Next, the
final period corresponds to the case when the dark disc has cleared
the caustic and the bright disc continues to overlap with the
caustic. During this period, the lightcurve reaches a peak that for
most of the parameter space is global and for the rest of the
parameter space local.  At the end of the period, the growth of the
magnification is negative infinite. the respective point appears as a
second saddle point on the lightcurve. Past this point the
magnification of any finite source will decay in roughly the same
manner.

@@

\begin{figure}
\centering
    \includegraphics[width = 0.48\textwidth]{figures/M87_shape.eps}
        \includegraphics[width = 0.48\textwidth]{figures/M87_lc.eps}
\caption{\label{fig:M87_plots} The one dimensional profile obtained
  from the numerical integration along the ordonata axis of the source
  image is presented in the upper panel. The corresponding
  lightcurve associated to the 1D profile is presented in the lower
  panel. The points $\rm p_2$ and $\rm p_3$ mark the start and 
  end of the equivalent
  dark disk and are associated with the moments of their overlap with
  the caustic: $\rm t_2$ and $\rm t_3$. $\rm h_{w}$ denotes the maximum width of the crescent.}
\end{figure}
